Untitled

### 2024-10-04 analysis/paper/access-ice.qmd
Adds extent plot

---
title: "Untitled"
format: pdf
execute: 
  echo: false
  warning: false
  message: false
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
---

@@

@@

{>>Why is ERA5 better or worse than satellites?<<}

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 era5 = ERA5(),
                 access = S2_reanalysis()
                )
```
```{r extent_daily}
extent_daily <- datasets |>
    lapply(extent) |>
    lapply(\(x) ReadNetCDF(x, "aice")) |>
    rbindlist(idcol = "model") |>
    _[, let(lon = NULL, lat = NULL)]

@@

extent_daily |>  
  copy() |>
  _[aice == 0, aice := NA] |> 
  # _[, daice := c(diff(aice), 0)] |>
  # _[, daice := daice - median(daice, na.rm = TRUE), by = .(model, yday(time))] |>
  # _[abs(daice) > 5e11 & model == "nsidc",
  #   let(daice = NA,
  #       aice = NA)] |>
  # _[, aice := frollmean(aice, 10, align = "center", na.rm = TRUE), by = model] |>
  _[!(month(time) == 2 & mday(time) == 29)] |> 
  _[, time := update(time, year = 2001)] |> 
  _[, median_ci(aice), by = .(time, model)]  |> 
  ggplot(aes(time, mid)) +
  geom_ribbon(data = \(x) x[model == "access"],
    aes(ymin = low, ymax = high, 
                  color = model, fill = model), alpha = 0.3) +
  geom_line(aes(color = model), linewidth = 1) +
#   scale_color_models +
#   scale_fill_models +
  scale_y_continuous("Extent",
                     labels = labels_extent) +
  scale_x_datetime(NULL, date_labels = "%b", date_breaks = "1 month")
```

@@

{>>Add reginoal extents?<<}

@@

@@

>>>>

---
title: "Untitled"
format: pdf
execute: 
  echo: false
  warning: false
  message: false
  cache: true
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
bibliography: references.bib
---

@@

[@comiso2023]

@@

{\>\>Why is ERA5 better or worse than satellites?\<\<}

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 era5 = ERA5(),
                 access = S2_reanalysis()
                )
```

```{r extent_daily}
extent_daily <- datasets |>
    lapply(extent) |>
    lapply(\(x) ReadNetCDF(x, "aice")) |>
    rbindlist(idcol = "model") |>
    _[, let(lon = NULL, lat = NULL)]

```{r}
extent_climatology <- extent_daily |>  
  copy() |>
  _[aice == 0, aice := NA] |> 
  _[!(month(time) == 2 & mday(time) == 29)] |> 
  _[, time := update(time, year = 2001)] |> 
  _[, median_ci(aice), by = .(time, model)]

@@

extent_climatology |>
  ggplot(aes(time, mid)) +
  geom_ribbon(data = \(x) x[model == "access"],
              aes(ymin = low, ymax = high, 
                  color = model, fill = model), alpha = 0.3) +
  geom_line(aes(color = model), linewidth = 1) +
#   scale_color_models +
#   scale_fill_models +
  scale_y_continuous("Extent",
                     labels = labels_extent) +
  scale_x_datetime(NULL, date_labels = "%b", date_breaks = "1 month")
```

@@

{\>\>Add reginoal extents?\<\<}

@@

```{r hindcast-extent}
rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_extent.Rds")),
      S2 = readRDS(here::here("data/derived/S2_hindcast_extent.Rds")), 
      idcol = "model") |>
    _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                                lag = as.numeric(as.Date(time) - forecast_time),
                                model)] |>
    _[, time := forecast_time + lag] |>
    _[, time2 := update(time, year = 2001)] |>
    _[, month := lubridate::month(forecast_time, label = TRUE)] |>
    # _[]
    ggplot(aes(time2, mid)) +
    geom_line(data = extent_climatology[model == "cdr"][, model := NULL], 
              aes(x = as.Date(time))) +    
    geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
    geom_point(data = \(x) x[lag == 1], aes(fill = month), 
               shape = 21, size = 3) +

    scale_x_date(NULL, date_breaks = "1 month", 
                 date_labels = "%b", expand = c(0, 0)) +
    scale_y_continuous(labels = labels_extent) +
    scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
                       aesthetics = c("fill", "color"), guide = "none") +
    facet_wrap(~ model, ncol = 1)
```

@@

# References

::: {#refs}
:::
